---
title: "Met Office Historic Station Data"
author: "Kate Belisle"
date: "2/10/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
require(tidyverse)
require(data.table) # %like%
require(lubridate)
require(bsts) # LastDayInMonth
knitr::opts_chunk$set(echo = TRUE)
```

```{r process-urls}
historical=c("https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/aberporthdata.txt","Aberporth",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/armaghdata.txt","Armagh",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/ballypatrickdata.txt","Ballypatrick Forest",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/bradforddata.txt","Bradford",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/braemardata.txt","Braemar",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/cambornedata.txt","Camborne",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/cambridgedata.txt","Cambridge NIAB",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/cardiffdata.txt","Cardiff Bute Park",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/chivenordata.txt","Chivenor",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/cwmystwythdata.txt","Cwmystwyth",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/dunstaffnagedata.txt","Dunstaffnage",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/durhamdata.txt","Durham",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/eastbournedata.txt","Eastbourne",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/eskdalemuirdata.txt","Eskdalemuir",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/heathrowdata.txt","Heathrow",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/hurndata.txt","Hurn",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/lerwickdata.txt","Lerwick",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/leucharsdata.txt","Leuchars",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/lowestoftdata.txt","Lowestoft",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/manstondata.txt","Manston",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/nairndata.txt","Nairn",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/newtonriggdata.txt","Newton Rigg",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/oxforddata.txt","Oxford",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/paisleydata.txt","Paisley",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/ringwaydata.txt","Ringway",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/rossonwyedata.txt","Ross-on-Wye",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/shawburydata.txt","Shawbury",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/sheffielddata.txt","Sheffield",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/southamptondata.txt","Southampton",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/stornowaydata.txt","Stornoway Airport",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/suttonboningtondata.txt","Sutton Bonington",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/tireedata.txt","Tiree",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/valleydata.txt","Valley",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/waddingtondata.txt","Waddington",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/whitbydata.txt","Whitby",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/wickairportdata.txt","Wick Airport",
            "https://www.metoffice.gov.uk/pub/data/weather/uk/climate/stationdata/yeoviltondata.txt","Yeovilton")
n=length(historical)
hstations=tibble(url=historical[seq(1,n,2)],
station=historical[seq(2,n,2)])
hstations$fname=basename(hstations$url)
n=n/2
for(i in seq(1,n)){
#    with(hstations,download.file(url[i], fname[i]))
}
```

##Now process a sample data set.

Yeovilton  
Location: 355100E 123200N, Lat 51.006 Lon -2.641, 20 metres amsl  
Estimated data is marked with a * after the value.  
Missing data (more than 2 days missing in month) is marked by  ---.  
Sunshine data taken from an automatic Kipp & Zonen sensor marked with a #, otherwise sunshine data taken from a Campbell Stokes recorder.  
   yyyy  mm   tmax    tmin      af    rain     sun  
              degC    degC    days      mm   hours  
   1964   9   20.5     8.8       0    37.4     ---
   1964  10   13.6     4.2       5    77.8     ---
*snip*              

First row is name, quality check
Process Location
Not sure if tmax is mean(max) ?? I guess it is
af is air frost days
need to calculate days in month

```{r read-data-fcn}
readMet<-function(station,fname){
  print(station)
  indat<-read_lines(fname,skip_empty_rows = TRUE)#,n_max=1,skip = 2)
  return(indat)
}

```

```{r process-stations}
locations<-c()
metdata<-NULL
cnames<-c("myear",
             "mmonth",
             "tmax",
             "tmin",
             "af",
             "rain",
             "sun")

ctypes<-cols("myear"=col_number(),
             "mmonth"=col_number(),
             "tmax"=col_number(),
             "tmin"=col_number(),
             "af"=col_number(),
             "rain"=col_number(),
             "sun"=col_number())

for(i in 1:n){
  
  raw<- (with(hstations,readMet(station[i],paste0('metdata/',fname[i])  )))
  loc<-raw[raw %like% "Location:"]
  locations<-c(locations,loc)
  fdata<-which(raw %like% "degC")+1
#  print(raw[fdata])
  ldata<-which(raw %like% "Provisional")
  if (length(ldata)>0)ldata=ldata[1]
  if (length(ldata)==0)ldata=length(raw)+1
  rmetdata<-read_delim(raw[(fdata):(ldata-1)],
                       delim=" ",col_names=cnames,
                       col_types =   ctypes,
                       trim_ws=TRUE,na="---") %>%
            mutate(station=hstations$station[i],
                    tavg=((tmax)+(tmin))/2,
                    trng=((tmax)-(tmin)),
                   mdate=(make_date(myear,mmonth,1))) 
  print(tail(rmetdata))
#  problems(rmetdata)
  metdata<-bind_rows(metdata,rmetdata)
  # loc<- str_split(with(hstations,readMetLoc(names[i], paste0('metdata/',fname[i])  ) ),",")[[1]]
  # lat=as.numeric(str_split(loc[2]," ")[[1]][3])    
  # lon=as.numeric(str_split(loc[2]," ")[[1]][5]) 
  # amsl=as.numeric(str_split(loc[3]," ")[[1]][1]) 
  # locations<-c(locations,list(lat,lon,amsl))
}
saveRDS(metdata,'metdata.rds')
readr::write_lines(locations,"locations.txt")
print(locations)

```

